\normalfont\spacedallcaps{Hazardous asteroids forecast via Markov random fields}

### 2021-11-25 Manuscript/manuscript.tex
PUB

@@

In this section we are going to review the theoretical concepts that underlies to the probabilistic methods here used:  we will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller et al. \cite{koller2009probabilistic}, H{\o}jsgaardand \cite{hojsgaard2012graphical} et al.  and Russel et al. \cite{russell2010artificial}.  Furthermore we will provide also a rapid overview of the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches,  since we used some of its concepts in the preliminary analysis of the dataset. Finally we concluded this chapter with a rapid overview about the concept of algorithm intepretability, which together with the accuracy and the confusion matrix, is necessary for the assessment of a ML performances. On the other side the concepts related to the celestial mechanics here used will be described, following the Murray approach \cite{murray1999solar} into the Appendix A

@@

\chapter{Results}

@@

>>>>

The description and the forecast of physical \footnote{Where here physical should be inteded as mesuarable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms one can formulate theoretical models that provide equations that describe such phenomena. Such approach, which will be called here \textit{ab-initio}. On the other side one can start from data and fit them into some general models (with fixed or not fixed number of parameters). This method will be called here as machine learning (ML) method. In the first case one have the advantage to have an almost full explanation of the meaning of the laws obtained and an overall idea of why nature works in this way. The toll that has to be paid in this case is that, since the phenomena are much more complicated with respect to the starting principles, a consistent part of them may be excluded by the assumption made at the beginning. Furthermore the calculation of the solution given the dynamical equation can be computationally expensive\footnote{This is the case for instance of Hartree-Fock method or of Density Functional Theory: such methods rewrite, with some approximations, the Schroedinger equation in to a much more computationally affordable way. However also in this case the solution for solids can be very expensive \cite{martin_2004}} . On the other side, the ML methods since they make much more general assumptions and work directly on data, are not limited to a particular class of phenomena. Therefore they do not require a development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation. The toll that has to be paid in this case it that the underlying mechanism by which a forecast is preferred to another one may be not so interpretable. Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century astronomers know that mercury follows an elliptical orbit that instead of being fixed rotates around the Sun as shown in Fig. \ref{Perihelion_precession2} Such phenomenon is called \textit{Precession of the perihelion of Mercury}. The crucial point is that there was no way to explain from the Newton gravity theory such phenomenon. At this point scholars had two paths: describe its motion with an empirical law (perhaps with an empirical modification of Newtons gravity law) or instead reformulate from scratch the Newton gravity. The second approach was the one followed by A.Einstein with his general relativity theory (further details are given in the A.Zee textbook \cite{zee2013einstein}). The graphical methods represent a good compromise  between these two approaches: they can be used for every physical phenomena and, since they provide the conditional dependencies/independences they also give a fine and intuitive representation  of their mechanisms. Therefore, if the theory of the process is known, one can use this class of algorithm and compare its finding with the theory. In this way one can not only evaluate the quality of the forecasts, but also of the model provided by the graphical methods. The aim of the present work is to perform such analysis for a dataset that contain the features and the hazardousness \footnote{For Earth} of asteroids. For this dataset, indeed, the relationships between the various features are known. The present work is organized as follows: first a brief introduction about the basic theoretical concepts of the statistical methods here used will be provided, then the dataset will be described and finally, the main results here obtained with graphical methods will be presented as compared with other type of machine learning algorithms. Furthermore an appendix about the theoretical concepts of the theory that explains and interconnects the dataset features is also provided at the end of the present work.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{Figures/Perihelion_precession2.png}
\caption{Precession of the perihelion of Mercury. Image taken from \cite{Perihelion_precession}}
\label{Perihelion_precession2}
\end{center}
\end{figure}

@@

In this section we are going to review the theoretical concepts on which the probabilistic methods here used are based. We will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller et al. \cite{koller2009probabilistic}, H{\o}jsgaardand \cite{hojsgaard2012graphical} et al.  and Russel et al. \cite{russell2010artificial}.  Furthermore we will provide also a rapid overview of the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches,  since we used some of its concepts in the preliminary analysis of the dataset. Finally we concluded this chapter with a rapid overview about the concept of algorithm intepretability, which together with the accuracy and the confusion matrix, is necessary for the assessment of a ML performances. On the other side the concepts related to the celestial mechanics here used will be described, following the Murray approach \cite{murray1999solar} into the Appendix A

@@

\chapter{Results and discussion}

@@

The Glasso (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that correctly capture the physical connections of the dataset feature. In particular for both models we saw that the no one of the orbital parameters are conditionally dependent with respect to the diameter of the asteroid, as stated by theory. Furthermore the mgm model is able to provide forecasts about the hazardousness of the asteroids with performances that are almost equal or better  than the logistic and QDA algorithms. On the other side the fact that its performances are lower with respect to the RF and SVM is definitely compensated by its interpretability.
